### 2024-04-01 paper/paper.qmd
Updated title page and changed file names.

---
title: "The Analysis of MLB Team Statistics between 2012 - 2018 on Predicting Games Won per Season ^[Code and data are available at: https://github.com/YcartXin/Analysis-of-Men-in-Saudi-Arabia-on-WWOH; Reproduction DOI: https://doi.org/10.48152/ssrp-1wa4-dk30]"
author: 
  - Tracy Yang
date: today
date-format: long
abstract: "hehe."
format: 
  typst:
    toc: true
number-sections: true
bibliography: references.bib
---
```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....
We examine the Major League Baseball team data from 2012 to 2018 focusing on a team's batting average (AVG), defensive efficiency ratio (DER), and earned run average (ERA) to predict the number of games a team would win out of 162 in a season. We estimate an Bayesian hierarchical model which explains their attributes based on their functionality.

# Data {#sec-data}

Some of our data is of penguins , from .

Talk more about it.

And also planes. (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix ].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in .

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In 1 we implement a posterior predictive check. This shows...

In 2 we compare the posterior with the prior. This shows...

## Diagnostics

is a trace plot. It shows... This suggests...

 is a Rhat plot. It shows... This suggests...

\newpage

# References

### 2024-04-02 paper/paper.qmd
Developed introduction

@@
 ---
-title: "The Analysis of MLB Team Statistics between 2012 - 2018 on Predicting Games Won per Season ^[Code and data are available at: https://github.com/YcartXin/Analysis-of-Men-in-Saudi-Arabia-on-WWOH; Reproduction DOI: https://doi.org/10.48152/ssrp-1wa4-dk30]"
+title: "MLB Team Statistics between 2012 - 2018 on Predicting Games Won per Season ^[Code and data are available at: https://github.com/YcartXin/MLB_Team_Stats"
 author: 
@@
 
+Behind the entertaining personalities and eye-catching traditions of baseball is a game where teams and are meticulously crafted with calculation.
+Baseball's fascination with numbers persisted since its creation (Swhwarz, 2004).
+Runs scored have always been recorded to evaluate performance, while other statistics developed overtime into the complicated net of data that now reside upon the MLB official site ().
+However, a team's performance can differ greatly year to year.
+A team's last year wins are not often indicative of the coming season.
+Many factors play into the seasonal performance of a team including more unpredictable variables such as injuries, and luck.
+However, with the common narrative of baseball being a numbers game, we aim to estimate the impact of important statistics of a team on its games won per season, more specifically, the progressing 2024 season.
+
 You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
@@
 # Data {#sec-data}
-
-Some of our data is of penguins , from .
 

### 2024-04-03 paper/paper.qmd
Progressing on introduction

@@
 ---
-title: "MLB Team Statistics between 2012 - 2018 on Predicting Games Won per Season ^[Code and data are available at: https://github.com/YcartXin/MLB_Team_Stats"
+title: "MLB Team Statistics between 2000 - 2023 on Predicting Games Won per Season ^[Code and data are available at: https://github.com/YcartXin/MLB_Team_Stats]"
 author: 
@@
 date-format: long
-abstract: "hehe."
+abstract: "Using MLB key team statistics between 2000 and 2023, this paper investigates the outcome of the 2024 season using each team's current statistics on predicting their total wins out of 162 games.
+We also investigate the common conception that baseball is a numbers game by evaluating the predictability of wins that a team would obtain during a seasion from looking at their batting average (AVG), defensive efficiency ratio (DER), and earned run average (ERA)."
 format: 
@@
 Baseball's fascination with numbers persisted since its creation (Swhwarz, 2004).
-Runs scored have always been recorded to evaluate performance, while other statistics developed overtime into the complicated net of data that now reside upon the MLB official site ().
-However, a team's performance can differ greatly year to year.
+Runs scored have always been recorded to evaluate performance, while other statistics developed overtime into the complicated net of data that now reside upon the MLB official site (XX).
+The term "sabermetrics" even developed to describe in specific baseball statistics (XX) However, a team's performance can differ greatly year to year.
 A team's last year wins are not often indicative of the coming season.
@@
 
+Important statistics for a team in this paper is defined as the team's batting average (AVG), defensive efficiency ratio (DER), and earned run average (ERA).
+These three statistics each refer to an important aspect of a strong team: hitting, pitching, and fielding prospectively.
+The specifics of these indicators will be elaborated on in @sec-data.
+Many 2024 season predictions exist...
+
 You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
-
-The remainder of this paper is structured as follows. @sec-data....
-We examine the Major League Baseball team data from 2012 to 2018 focusing on a team's batting average (AVG), defensive efficiency ratio (DER), and earned run average (ERA) to predict the number of games a team would win out of 162 in a season.
-We estimate an Bayesian hierarchical model which explains their attributes based on their functionality.
 
@@
 
-\appendix
-
-# Appendix {-}
-
 # Additional data details

### 2024-04-05 paper/paper.qmd
Updated paper and simulated data

@@
 abstract: "Using MLB key team statistics between 2000 and 2023, this paper investigates the outcome of the 2024 season using each team's current statistics on predicting their total wins out of 162 games.
-We also investigate the common conception that baseball is a numbers game by evaluating the predictability of wins that a team would obtain during a seasion from looking at their batting average (AVG), defensive efficiency ratio (DER), and earned run average (ERA)."
+We also investigate the common conception that baseball is a numbers game by evaluating the predictability of wins that a team would obtain during a seasion from looking at their batting average (AVG), defensive efficiency ratio (DffEff), and earned run average (ERA)."
 format: 
@@
 
+\newpage
+
 # Introduction
@@
 
-Important statistics for a team in this paper is defined as the team's batting average (AVG), defensive efficiency ratio (DER), and earned run average (ERA).
+Important statistics for a team in this paper is defined as the team's batting average (AVG), defensive efficiency ratio (DffEff), and earned run average (ERA).
 These three statistics each refer to an important aspect of a strong team: hitting, pitching, and fielding prospectively.
 The specifics of these indicators will be elaborated on in @sec-data.
-Many 2024 season predictions exist...
+Among present research, 2024 season predictions exist, however, they mainly focus on the win for a game or series for betting purposes.
+This paper is unique in its approach to predict and investigates the predictability of overall seasonal wins for baseball teams.
+
+The rest of the paper is structured as follows.
+@sec-data (Data) will explain the source of the data used, its collection process, and the process of cleaning that led to the final data used as well as potential biases.
+@sec-model...
+
+The overall analysis shows...
 
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-## Second discussion point
-
-## Third discussion point
-
 ## Weaknesses and next steps

### 2024-04-05 scripts/00-simulate_data.R

@@ -1,21 +1,47 @@
 #### Preamble ####
-# Purpose: Simulates an analysis of selected pokemon's weaknesses and strengths
+# Purpose: Simulates the data sets of MLB with variables of interest (AVG, DffEff, ERA, W).
 # Author: Tracy Yang
-# Date: 11 February 2023
+# Date: 2nd April 2023
 # Contact: ycart.yang@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: -
-# Any other information needed?
-
 
 #### Workspace setup ####
 library(tidyverse)
+library(dplyr)
 
 
 #### Simulate data ####
+set.seed(11)
+teams <- c("NYY", "BOS", "BAL", "TB", "TOR", 
+          "DET", "CLE", "MIN", "KC", "CWS",
+          "LAA", "TEX", "SEA", "HOU", "OAK",
+          "ATL", "WSH", "PHI", "NYM", "MIA",
+          "PIT", "MIL", "CHC", "CIN", "STL",
+          "LAD", "AZ", "SD", "SF", "COL")
 
-data <- 
+#Data simulated for one season (2000)
+data_00 <- 
   tibble(
-    level = rnorm(n = 800, mean = 100, sd = 10) |> floor(),
-    weakness = sample(x = c("water", "ground", "fire"), size = 800, replace = TRUE)
+    Year = 2000,
+    Team = sample(x = teams, size = 30, replace = FALSE),
+    AVG = rnorm(n = 30, mean = 0.2, sd = 0.1),
+    ERA = rnorm(n = 30, mean = 4, sd = 1),
+    DffEff = rnorm(n = 30, mean = 0.7, sd = 0.01)
   )
+
+#Data simulated for one season (2001)
+data_01 <- 
+  tibble(
+    Year = 2001,
+    Team = sample(x = teams, size = 30, replace = FALSE),
+    AVG = rnorm(n = 30, mean = 0.2, sd = 0.1),
+    ERA = rnorm(n = 30, mean = 4, sd = 1),
+    DffEff = rnorm(n = 30, mean = 0.7, sd = 0.01)
+  )
+
+#Simulating combining data for different seasons
+sim_data <- rbind(data_00, data_01)
+
+#Testing simulated data
+XX

### 2024-04-05 scripts/01-download_data.R

@@ -9,18 +9,9 @@
 
 
 #### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
-
-
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
-
-         
+# Team data downloaded from https://www.baseball-reference.com/leagues/.
+# Hitting, pitching, and fielding data downloaded separately for each year to be combined later.
+# XX - XX data were downloaded, in total XX files.

### 2024-04-08 scripts/00-simulate_data.R
Finished simulation and in progress download data

@@ -44,4 +44,4 @@ data_01 <-
 sim_data <- rbind(data_00, data_01)
 
 #Testing simulated data
-XX
+

### 2024-04-08 scripts/01-download_data.R

@@ -1,15 +1,13 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from Baseball Reference
+# Author: Tracy Yang
+# Date: 6th April 2024
+# Contact: ycart.yang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: -
 
 
 #### Workspace setup ####
-
 
 #### Download data ####
 # Team data downloaded from https://www.baseball-reference.com/leagues/.

### 2024-04-08 paper/paper.qmd
Added tests for simulated data, added more data frames.

@@
 Important statistics for a team in this paper is defined as the team's batting average (AVG), defensive efficiency ratio (DffEff), and earned run average (ERA).
-These three statistics each refer to an important aspect of a strong team: hitting, pitching, and fielding prospectively.
+These three statistics each refer to an important aspect of a strong team: hitting, pitching, and fielding prospectively, hence important indicators on the wins a team is predicted to earn in a season.
 The specifics of these indicators will be elaborated on in @sec-data.
@@
 The rest of the paper is structured as follows.
-@sec-data (Data) will explain the source of the data used, its collection process, and the process of cleaning that led to the final data used as well as potential biases.
+@sec-data (Data) will explain the source of the data used, its collection process, and the process of cleaning that led to the final data used.
+The data section will also explore the potential biases the collected and final data may possess.
 @sec-model...
@@
 The overall analysis shows...
-
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
 

### 2024-04-08 scripts/00-simulate_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulates the data sets of MLB with variables of interest (AVG, DffEff, ERA, W).
 # Author: Tracy Yang
-# Date: 2nd April 2023
+# Date: 2 April 2023
 # Contact: ycart.yang@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: -

### 2024-04-08 scripts/01-download_data.R

@@ -1,13 +1,11 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from Baseball Reference
+# Purpose: Downloads and saves the data from Baseball Reference.
 # Author: Tracy Yang
-# Date: 6th April 2024
+# Date: 6 April 2024
 # Contact: ycart.yang@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: -
 
-
-#### Workspace setup ####
 
 #### Download data ####
 # Team data downloaded from https://www.baseball-reference.com/leagues/.

### 2024-04-10 scripts/00-simulate_data.R
Updated csv files and finished half of the data cleaning process.

@@ -27,7 +27,8 @@ data_00 <-
     Team = sample(x = teams, size = 30, replace = FALSE),
     AVG = rnorm(n = 30, mean = 0.2, sd = 0.1),
     ERA = rnorm(n = 30, mean = 4, sd = 1),
-    DffEff = rnorm(n = 30, mean = 0.7, sd = 0.01)
+    DffEff = rnorm(n = 30, mean = 0.7, sd = 0.01),
+    W = round(rnorm(n = 30, mean = 82, sd = 3))
   )
 
 #Data simulated for one season (2001)
@@ -37,11 +38,25 @@ data_01 <-
     Team = sample(x = teams, size = 30, replace = FALSE),
     AVG = rnorm(n = 30, mean = 0.2, sd = 0.1),
     ERA = rnorm(n = 30, mean = 4, sd = 1),
-    DffEff = rnorm(n = 30, mean = 0.7, sd = 0.01)
+    DffEff = rnorm(n = 30, mean = 0.7, sd = 0.01),
+    W = round(rnorm(n = 30, mean = 82, sd = 3))
   )
 
 #Simulating combining data for different seasons
 sim_data <- rbind(data_00, data_01)
 
 #Testing simulated data
+sim_data$Year |> unique() == c(2000, 2001)
+sim_data$Year |> length() == 60
+sim_data$Team |> unique() |> length() == 30
+sim_data$Team |> is.character()
+sim_data$AVG |> is.numeric()
+sim_data$AVG |> min() >= 0
+sim_data$ERA |> is.numeric()
+sim_data$ERA |> min() >= 0
+sim_data$DffEff |> is.numeric()
+sim_data$DffEff |> max() <= 1
+sim_data$W |> is.numeric()
+
+
 

### 2024-04-10 scripts/01-download_data.R

@@ -8,6 +8,7 @@
 
 
 #### Download data ####
-# Team data downloaded from https://www.baseball-reference.com/leagues/.
+# Team data downloaded from https://www.baseball-reference.com/leagues/ and then 
+# converted from xls files to csv files to be saved in the raw data folder.
 # Hitting, pitching, and fielding data downloaded separately for each year to be combined later.
 # XX - XX data were downloaded, in total XX files.

### 2024-04-10 scripts/02-data_cleaning.R

@@ -1,44 +1,39 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw Baseball Reference batting, pitching, and fielding data from years 2000 to 
+# Author: Tracy Yang
+# Date: 8 April, 20204
+# Contact: ycart.yang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: -
+# Data source and preparation specified in 01-download_data and README.
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+#2000
+b2000 <- read_csv("data/raw_data/2000B.csv")
[…]

### 2024-04-13 paper/paper.qmd
Finished downloading, cleaning, and testing data. Data section in progress.

@@
 abstract: "Using MLB key team statistics between 2000 and 2023, this paper investigates the outcome of the 2024 season using each team's current statistics on predicting their total wins out of 162 games.
-We also investigate the common conception that baseball is a numbers game by evaluating the predictability of wins that a team would obtain during a seasion from looking at their batting average (AVG), defensive efficiency ratio (DffEff), and earned run average (ERA)."
+We also investigate the common conception that baseball is a numbers game by evaluating the predictability of wins that a team would obtain during a seasion from looking at their batting average (AVG), defensive efficiency ratio (DefEff), and earned run average (ERA)."
 format: 
@@
 
-Important statistics for a team in this paper is defined as the team's batting average (AVG), defensive efficiency ratio (DffEff), and earned run average (ERA).
-These three statistics each refer to an important aspect of a strong team: hitting, pitching, and fielding prospectively, hence important indicators on the wins a team is predicted to earn in a season.
+Important statistics for a team in this paper is defined as the team's batting average (AVG), defensive efficiency ratio (DefEff), and earned run average (ERA).
+These three statistics each refer to an important aspect of a strong team: hitting, pitching, and fielding prospectively.
+A baseball team cannot succeed with a weak end in any of the three categories.
+Therefore, these are important indicators on the estimand - the wins a team would achieve by the end of the 2024 seasons among 162 games.
 The specifics of these indicators will be elaborated on in @sec-data.
-Among present research, 2024 season predictions exist, however, they mainly focus on the win for a game or series for betting purposes.
+Among present research, 2024 season predictions exist, however, they mainly focus on the win for a specific game or series for betting purposes, focusing on the player data.
 This paper is unique in its approach to predict and investigates the predictability of overall seasonal wins for baseball teams.
@@
 
-Talk more about it.
-
-And also planes.
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+The data used in this analysis of different statistics on MLB Teams are sourced from Sports Reference, more specifically, Baseball Reference ().
+The data available on Baseball Reference is gathered by many contributors through a variety of sources such as the Lahmen database, and much of older data is accessed from third-party websites such as Retrosheet ().
+Baseball Reference stores and manipulate the data gathered using MySQL ().
+The data sets used in this investigation focus on the team statistics of the 30 teams in MLB and span from 2000 to 2010.
+The data sets were downloaded separately for each year among the batting, pitching, and fielding data.
+They were then combined into the final data set during the cleaning process.
+This analysis will be carried out in **R** @R using packages **tidyverse** @tidyverse, **dplyr** , **ggplot2** @ggplot2, **knitr** , **here **  XX.
 

### 2024-04-13 scripts/00-simulate_data.R

@@ -38,7 +38,7 @@ data_01 <-
     Team = sample(x = teams, size = 30, replace = FALSE),
     AVG = rnorm(n = 30, mean = 0.2, sd = 0.1),
     ERA = rnorm(n = 30, mean = 4, sd = 1),
-    DffEff = rnorm(n = 30, mean = 0.7, sd = 0.01),
+    DefEff = rnorm(n = 30, mean = 0.7, sd = 0.01),
     W = round(rnorm(n = 30, mean = 82, sd = 3))
   )
 
@@ -54,8 +54,8 @@ sim_data$AVG |> is.numeric()
 sim_data$AVG |> min() >= 0
 sim_data$ERA |> is.numeric()
 sim_data$ERA |> min() >= 0
-sim_data$DffEff |> is.numeric()
-sim_data$DffEff |> max() <= 1
+sim_data$DefEff |> is.numeric()
+sim_data$DefEff |> max() <= 1
 sim_data$W |> is.numeric()
 
 

### 2024-04-13 scripts/02-data_cleaning.R

@@ -9,9 +9,12 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(dplyr)
+library(janitor)
+
 
 #### Clean data ####
-
+## Preparing yearly data ##
 #2000
 b2000 <- read_csv("data/raw_data/2000B.csv")
 f2000 <- read_csv("data/raw_data/2000F.csv")
@@ -32,8 +35,70 @@ b2003 <- read_csv("data/raw_data/2003B.csv")
 f2003 <- read_csv("data/raw_data/2003F.csv")
 p2003 <- read_csv("data/raw_data/2003P.csv")
 data2003 <- cbind(b2003[1:30, c("Tm", "BA")], f2003[1:30, c("DefEff")], p2003[1:30, c("ERA", "W")])
+#2004
+b2004 <- read_csv("data/raw_data/2004B.csv")
+f2004 <- read_csv("data/raw_data/2004F.csv")
+p2004 <- read_csv("data/raw_data/2004P.csv")
+data2004 <- cbind(b2004[1:30, c("Tm", "BA")], f2004[1:30, c("DefEff")], p2004[1:30, c("ERA", "W")])
+#2005
+b2005 <- read_csv("data/raw_data/2005B.csv")
+f2005 <- read_csv("data/raw_data/2005F.csv")
+p2005 <- read_csv("data/raw_data/2005P.csv")
+data2005 <- cbind(b2005[1:30, c("Tm", "BA")], f2005[1:30, c("DefEff")], p2005[1:30, c("ERA", "W")])
+#2006
+b2006 <- read_csv("data/raw_data/2006B.csv")
+f2006 <- read_csv("data/raw_data/2006F.csv")
+p2006 <- read_csv("data/raw_data/2006P.csv")
+data2006 <- cbind(b2006[1:30, c("Tm", "BA")], f2006[1:30, c("DefEff")], p2006[1:30, c("ERA", "W")])
+#2007
+b2007 <- read_csv("data/raw_data/2007B.csv")
+f2007 <- read_csv("data/raw_data/2007F.csv")
+p2007 <- read_csv("data/raw_data/2007P.csv")
+data2007 <- cbind(b2007[1:30, c("Tm", "BA")], f2007[1:30, c("DefEff")], p2007[1:30, c("ERA", "W")])
+#2008
+b2008 <- read_csv("data/raw_data/2008B.csv")
+f2008 <- read_csv("data/raw_data/2008F.csv")
+p2008 <- read_csv("data/raw_data/2008P.csv")
[…]

### 2024-04-13 scripts/03-test_data.R

@@ -1,15 +1,40 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests the cleaned MLB data of teams between 2000 to 2010 inclusive.
+# Author: Tracy Yang
+# Date: 11 April, 2024
+# Contact: ycart.yang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 02-data_cleaning.R
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(here)
+
+#### Importing ####
+mlb_test <- read_csv("data/analysis_data/mlb_data.csv", show_col_types = FALSE)
+
 
 #### Test data ####
+
+## Test Team ##
+mlb_test$Team |> unique() |> length() == 30
+mlb_test$Team |> is.character()
+
+## Test AVG ##
+mlb_test$AVG |> is.numeric()
+mlb_test$AVG |> min() >= 0
+mlb_test$AVG |> max() <= 0.99
+
+## Test ERA ##
+mlb_test$ERA |> is.numeric()
+mlb_test$ERA |> min() >= 0
+
+## Test DefEff ##
+mlb_test$DefEff |> is.numeric()
+mlb_test$DefEff |> min >= 0
+mlb_test$DefEff |> max() <= 1
+
+## Test W ##
+mlb_test$W |> is.numeric()
+mlb_test$W |> length() == 330

### 2024-04-14 paper/paper.qmd
progressing on data section

@@
 date-format: long
-abstract: "Using MLB key team statistics between 2000 and 2023, this paper investigates the outcome of the 2024 season using each team's current statistics on predicting their total wins out of 162 games.
+abstract: "Using Major League Baseball key team statistics between 2000 and 2023, this paper investigates the outcome of the 2024 season using each team's current statistics on predicting their total wins out of 162 games.
 We also investigate the common conception that baseball is a numbers game by evaluating the predictability of wins that a team would obtain during a seasion from looking at their batting average (AVG), defensive efficiency ratio (DefEff), and earned run average (ERA)."
@@
 library(tidyverse)
+library(here)
+library(arrow)
+library(knitr)
+
+mlb_data <- read_parquet(here("data/analysis_data/mlb_data.parquet"))
+
 ```
@@
 Runs scored have always been recorded to evaluate performance, while other statistics developed overtime into the complicated net of data that now reside upon the MLB official site (XX).
-The term "sabermetrics" even developed to describe in specific baseball statistics (XX) However, a team's performance can differ greatly year to year.
+The term "sabermetrics" even developed to describe in specific baseball statistics (XX) However,in Major League Baseball, a team's performance can differ greatly year to year.
 A team's last year wins are not often indicative of the coming season.
-Many factors play into the seasonal performance of a team including more unpredictable variables such as injuries, and luck.
-However, with the common narrative of baseball being a numbers game, we aim to estimate the impact of important statistics of a team on its games won per season, more specifically, the progressing 2024 season.
+Many factors play into the seasonal performance of a team including more unpredictable variables such as injuries and luck.
+However, with the common narrative of baseball being a numbers game, we aim to estimate the impact of important statistics of a team on its games won per season in MLB, more specifically, the progressing 2024 season.
 
@@
 
+\newpage
+
 # Data {#sec-data}
@@
 They were then combined into the final data set during the cleaning process.
-This analysis will be carried out in **R** @R using packages **tidyverse** @tidyverse, **dplyr** , **ggplot2** @ggplot2, **knitr** , **here **  XX.
+This cleaning and analysis are carried out in **R** @R using packages **tidyverse** @tidyverse, **dplyr** , **ggplot2** @ggplot2, **knitr** @knitr, **here ** @here, **dplyr** @dplyr, **janitor** @janitor, **arrow** @arrow  XX.
+
+## Measurement and Collection
+There are in total 33 raw data sets of MLB team statistics in the raw data folder for constructing the model, these are split into 11 years and 3 categories (batting, pitching, fielding) for each team of every year from 2000 to 2010.
+Among each category, one representative indicator is chosen to demonstrate the overall strength of that category and be included as a variable for predicting wins over a season.
+
+The batting data in a year contains 29 variables detailing the overall offensive strength of hitters in each team.
+These variables include average runs per game or 2nd base reached over a season.
+Among these variables, we are interested in the batting average (AVG).
+AVG is calculated by total hits divided by total at-bats.
+Hits is defined by the number of times a batter strikes the ball into fair territory and reaches base, in the absence of an error or a fielder's choice ().
+AVG is a good indicator for offensive capabilities as it minimizes the impact of the pitcher by discounting walks and it is one of the universal tools to measure a hitter's success.
+
+The pitching data in a year contains 36 variables for each team detailing the bullpen's strength in initiating outs and preventing runs.
+Among variables in the data, we are interested in the Earned Run Average (ERA) calculated by 9 x earned runs/innings pitched ().
+This number indicates the average earned runs allowed per 9 innings, which is better for a team if lower.
+This indicator tests the fundamental purpose of pitchers in the bullpen as their goal is to prevent runs of the opposing team and accurately portrays the pitching ability of a team.
+
+The fielding data for each year contains 16 variables that details the defensive abilities of a team to get the hitter out after a ball is in play.
+The variable chosen to represent this ability is Defensive Efficiency (DefEff), calculated by 1 - (H^[H = Hit] + ROE^[ROE = Reached on Error, when a batter reaches base due to a defensive error] - HR^[HR = Home Run]) / (PA^[PA = Plate Appearance] - BB^[BB = Base on Balls, better known as a walk which happens when the pitcher receives 4 balls from the umpire] - SO^[SO = Strike Out, when the batter is out of play from three strikes from the umpire] - HBP^[HBP = Hit By Pitch, when a batter is struck directly by a pitch without swinging and is awarded a walk] - HR), which is the percentage of balls in play converted into outs ().
+This indicator estimates the ability of the catcher and fielders to efficiently prevent runs after a ball allows hitters to reach base and home.
+
+## Data Exploration ##
+
+There are thirty data points per year as there are thirty teams in MLB.
+In the data used for this investigation, there are three necessary predictors: AVG, ERA, DefEff.
+Team names is necessary to identify teams throughout the years and the number of wins is imported in with the pitching data.
+XX shows a sample of the the cleaned data with 5 variables.
+```{r}
+#| echo: false
+#| label: tbl-sample
+#| tbl-cap: Sample of Cleaned MLB Prediction for Model Building
+mlb_data |> slice(1:5) |> kable()
+
+```
 

### 2024-04-14 scripts/02-data_cleaning.R

@@ -11,6 +11,7 @@
 library(tidyverse)
 library(dplyr)
 library(janitor)
+library(arrow)
 
 
 #### Clean data ####
@@ -73,7 +74,7 @@ data2010 <- cbind(b2010[1:30, c("Tm", "BA")], f2010[1:30, c("DefEff")], p2010[1:
 
 ## Combining yearly data ##
 mlb_10 <- rbind(data2000, data2001, data2002, data2003, data2004, data2005, 
-                data2006, data2007, data2008, data2009, data2010) |> clean_names()
+                data2006, data2007, data2008, data2009, data2010)
 
 ## Cleaning yearly data ##
 current <- c("Arizona Diamondbacks", "Atlanta Braves", "Baltimore Orioles", 
@@ -95,10 +96,13 @@ change <- c("AZ", "ATL", "BAL", "BOS", "CHC",
            "PHI", "PIT", "SD", "SF", "SEA",
            "STL", "TB", "TEX", "TOR", "WSH", 
            "WSH", "TB", "LAA", "LAA")
-mlb_10$tm <- ifelse(mlb_10$tm %in% current, 
-                  change[match(mlb_10$tm, current)], 
-                  mlb_10$tm)
+mlb_10 <- mlb_10 |>
+  mutate(Tm = case_when(
+    mlb_10$Tm %in% current ~ change[match(mlb_10$Tm, current)],
+    TRUE ~ Tm
+  ))
+
 mlb_data <- mlb_10
 
 #### Save data ####
-write_csv(mlb_data, "data/analysis_data/mlb_data.csv")
+write_parquet(mlb_data, "data/analysis_data/mlb_data.parquet")

### 2024-04-14 scripts/03-test_data.R

@@ -10,9 +10,10 @@
 #### Workspace setup ####
 library(tidyverse)
 library(here)
+library(arrow)
 
 #### Importing ####
-mlb_test <- read_csv("data/analysis_data/mlb_data.csv", show_col_types = FALSE)
+mlb_test <- read_parquet("data/analysis_data/mlb_data.parquet")
 
 
 #### Test data ####

### 2024-04-16 paper/paper.qmd
Progressing on data section

@@
 The data used in this analysis of different statistics on MLB Teams are sourced from Sports Reference, more specifically, Baseball Reference ().
-The data available on Baseball Reference is gathered by many contributors through a variety of sources such as the Lahmen database, and much of older data is accessed from third-party websites such as Retrosheet ().
-Baseball Reference stores and manipulate the data gathered using MySQL ().
 The data sets used in this investigation focus on the team statistics of the 30 teams in MLB and span from 2000 to 2010.
@@
 ## Measurement and Collection
+The data available on Baseball Reference is gathered by many contributors through a variety of sources such as the Lahmen database, and much of older data is accessed from third-party websites such as Retrosheet ().
+Baseball Reference stores and manipulate the data gathered using MySQL ().
+A potential weakness in the data collection includes missing data for certain teams, especially for early seasons in the 19th century when statistics were not kept in database and are easily lost.
+However, this analysis uses recent season data between 2000 and 2010 which mitigates this concern.
+
 There are in total 33 raw data sets of MLB team statistics in the raw data folder for constructing the model, these are split into 11 years and 3 categories (batting, pitching, fielding) for each team of every year from 2000 to 2010.
@@
 These variables include average runs per game or 2nd base reached over a season.
-Among these variables, we are interested in the batting average (AVG).
+Among these variables, we are interested in the batting average (**AVG**).
 AVG is calculated by total hits divided by total at-bats.
@@
 The pitching data in a year contains 36 variables for each team detailing the bullpen's strength in initiating outs and preventing runs.
-Among variables in the data, we are interested in the Earned Run Average (ERA) calculated by 9 x earned runs/innings pitched ().
+Among variables in the data, we are interested in the Earned Run Average (**ERA**) calculated by 9 x earned runs/innings pitched ().
 This number indicates the average earned runs allowed per 9 innings, which is better for a team if lower.
@@
 The fielding data for each year contains 16 variables that details the defensive abilities of a team to get the hitter out after a ball is in play.
-The variable chosen to represent this ability is Defensive Efficiency (DefEff), calculated by 1 - (H^[H = Hit] + ROE^[ROE = Reached on Error, when a batter reaches base due to a defensive error] - HR^[HR = Home Run]) / (PA^[PA = Plate Appearance] - BB^[BB = Base on Balls, better known as a walk which happens when the pitcher receives 4 balls from the umpire] - SO^[SO = Strike Out, when the batter is out of play from three strikes from the umpire] - HBP^[HBP = Hit By Pitch, when a batter is struck directly by a pitch without swinging and is awarded a walk] - HR), which is the percentage of balls in play converted into outs ().
+The variable chosen to represent this ability is Defensive Efficiency (**DefEff**), calculated by 1 - (H^[H = Hit] + ROE^[ROE = Reached on Error, when a batter reaches base due to a defensive error] - HR^[HR = Home Run]) / (PA^[PA = Plate Appearance] - BB^[BB = Base on Balls, better known as a walk which happens when the pitcher receives 4 balls from the umpire] - SO^[SO = Strike Out, when the batter is out of play from three strikes from the umpire] - HBP^[HBP = Hit By Pitch, when a batter is struck directly by a pitch without swinging and is awarded a walk] - HR), which is the percentage of balls in play converted into outs ().
 This indicator estimates the ability of the catcher and fielders to efficiently prevent runs after a ball allows hitters to reach base and home.
 
-## Data Exploration ##
+## Cleaning and Exploration ##
+
+The cleaned data now contains 330 observations of 5 variables:
+
+- Team
+- Batting Average - hits over at bats
+- Defensive Efficiency - out rate after a ball is in play
+- Earned Run Average - average earned runs allowed per none innings
+- Wins
 
 There are thirty data points per year as there are thirty teams in MLB.
-In the data used for this investigation, there are three necessary predictors: AVG, ERA, DefEff.
+The predictor variables are batting average, earned average, and defensive efficiency.
 Team names is necessary to identify teams throughout the years and the number of wins is imported in with the pitching data.
-XX shows a sample of the the cleaned data with 5 variables.
+@tbl-sample shows a sample of the the cleaned data with 5 variables.
 ```{r}
@@
 #| label: tbl-sample
-#| tbl-cap: Sample of Cleaned MLB Prediction for Model Building
+#| tbl-cap: Sample of Cleaned MLB Prediction Data for Model Building
 mlb_data |> slice(1:5) |> kable()
 
+```
+
+We can understand this data better by understanding the distribution of some of the variables as well as the summary statistics.
+For instance, we can understand the competitiveness of MLB by seeing the distribution of wins per season, whether the teams differ significantly by wins can help us analyze whether there is a significant disparity in the strengths of different teams.
+@fig-wins shows the distribution of the wins data of each team among 162 games in a season between 2000 to 2010.
+
+```{r}
+#| echo: false
+#| label: fig-wins
+#| fig-cap: Distribution of Wins of each Team in a MLB Season between 2000 - 2010
+
+mlb_data |>
+  ggplot(aes(x = W)) +
+  geom_histogram(binwidth =  5, fill = "steelblue", alpha = 0.9, color = "white") +
+  theme_minimal() +
+  labs(
+    x = "Wins",
+    y = "Number of occurrences"
+  )
+
+```
+We can also get a sense of the average strength of the a MLB team in terms of the predictors by looking at their summary data. This is shown in @tbl-summary.
+
+```{r}
+#| echo: false
+#| label: tbl-summary
+#| tbl-cap: Summary of Cleaned MLB Prediction Data for Model Building
+summary_stat <- mlb_data |>
+  summarize(
+    Statistic = "Mean",
+    BA = mean(BA, na.rm =TRUE),
+    DefEff = mean(DefEff, na.rm = TRUE),
+    ERA = mean(ERA, na.rm = TRUE))
+
+summary_stat <- summary_stat |> add_row(
+  Statistic = "Median",
+  BA = median(mlb_data$BA), 
+  DefEff = median(mlb_data$DefEff), 
+  ERA = median(mlb_data$ERA))
+
+summary_stat |> kable()
 ```
@@
 
-# Additional data details
-
 # Model details {#sec-model-details}

### 2024-04-21 paper/paper.qmd
Progressing on model, added new data

@@
 ---
-title: "MLB Team Statistics between 2000 - 2023 on Predicting Games Won per Season ^[Code and data are available at: https://github.com/YcartXin/MLB_Team_Stats]"
+title: "MLB Team Statistics between 2000 - 2010 on Predicting Games Won per Season ^[Code and data are available at: https://github.com/YcartXin/MLB_Team_Stats]"
 author: 
@@
 date-format: long
-abstract: "Using Major League Baseball key team statistics between 2000 and 2023, this paper investigates the outcome of the 2024 season using each team's current statistics on predicting their total wins out of 162 games.
+abstract: "Using Major League Baseball key team statistics between 2000 and 2010, this paper investigates the outcome of the 2024 season using each team's current statistics on predicting their total wins out of 162 games.
 We also investigate the common conception that baseball is a numbers game by evaluating the predictability of wins that a team would obtain during a seasion from looking at their batting average (AVG), defensive efficiency ratio (DefEff), and earned run average (ERA)."
@@
 
+#install.packages("modelsummary")
+
 library(tidyverse)
@@
 library(knitr)
+library(modelsummary)
 
@@
 They were then combined into the final data set during the cleaning process.
-This cleaning and analysis are carried out in **R** @R using packages **tidyverse** @tidyverse, **dplyr** , **ggplot2** @ggplot2, **knitr** @knitr, **here ** @here, **dplyr** @dplyr, **janitor** @janitor, **arrow** @arrow  XX.
+The analysis in this paper will be carried out in **R** @R using packages **tidyverse** @tidyverse, **dplyr** , **ggplot2** @ggplot2, **knitr** @knitr, **here** @here, **arrow** @arrow  XX.
 
@@
 However, this analysis uses recent season data between 2000 and 2010 which mitigates this concern.
+Years 2000 to 2010 were chosen as this decade avoids the recent disturbances caused by COVID-19 yet is still immediate enough to be good training data for the model to predict the current MLB season.
 
@@
 
-- Team
-- Batting Average - hits over at bats
-- Defensive Efficiency - out rate after a ball is in play
-- Earned Run Average - average earned runs allowed per none innings
-- Wins
+- Team (Tm)
+- Batting Average (BA) - hits over at bats
+- Defensive Efficiency (DefEff) - out rate after a ball is in play
+- Earned Run Average (ERA) - average earned runs allowed per none innings
+- Wins (w)
 
@@
 #| tbl-cap: Sample of Cleaned MLB Prediction Data for Model Building
-mlb_data |> slice(1:5) |> kable()
+mlb_data |> slice(1:8) |> kable()
 
@@
 
-We can understand this data better by understanding the distribution of some of the variables as well as the summary statistics.
+We can understand this data better by understanding the distribution of some of the variables.
 For instance, we can understand the competitiveness of MLB by seeing the distribution of wins per season, whether the teams differ significantly by wins can help us analyze whether there is a significant disparity in the strengths of different teams.
@@
 ```
-We can also get a sense of the average strength of the a MLB team in terms of the predictors by looking at their summary data. This is shown in @tbl-summary.
+With 162 games per season, 50% win rate would count to 81 wins.
+However, @fig-wins shows that the highest frequencies lie around 85 to 90.
+This means that the distribution of wins over the 11 years is slightly skewed to the left.The teams in the league are generally quite competitive and close in wins, though, there are some weaker teams that stray further from 50% win rate.
+
+The data can be further explored with summary statistics.
+We can get a sense of the average strength of the a MLB team in terms of the three predictors by looking at their summary data.
+This is shown in @tbl-summary.
 
@@
 summary_stat |> kable()
+
 ```
+The mean and median of batting average are very close at around 0.264.
+This means that on average between 2000 to 2010, a MLB batter has 26.4% probability of getting a hit when at bat.
+The defensive efficiency is once again very similar between mean and median at around 0.690.
+MLB teams' defense against against a ball in play is quite strong, with around 69% of batters end up as outs when a ball is in play.
+However, the mean of earned run average is slightly higher than the median at around 4.39 compared to 4.35.
+This shows that there are some pitchers with very low ERAs that skew the mean to be lower.
+An average MLB pitcher between 2000 and 2010 overall allows around 4.39 runs per 9 innings.
+
+\newpage
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+There are two steps to the use of modeling in this paper.
+This paper will be utilizing a multiple linear regression containing three predictor variables to model their effects on the number of wins out of 162 for a MLB team.
+Subsequently, the model will be used to inference the team with the most wins in MLB in the 2024 season with their current team stats on batting average, defensive efficiency, and earned run average.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix ].
+@sec-setup @sec-justification will expand on the Bayesian analysis model and details of the multiple linear regression.
 
-## Model set-up
+## Model set-up {#sec-setup}
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Define $y_i$ as the number of wins a team would have by the end of 162 games.
+Subsequently, $\beta_0$ is the coefficient for the intercept and $\beta_1, \beta_2, \beta_3$ are respectively defined as the effect that batting average, defensive efficiency, and earned run average has on $y_i$ to be estimated.
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+$y_i|\mu_i, \sigma \sim \mbox{Normal}(\mu_i, \sigma)$\
+$\mu_i = \beta_0 + \beta_1\mbox{BA}_i + \beta_2\mbox{DefEff}_i + \beta_3\mbox{ERA}_i$\
+$\beta_0 \sim \mbox{Normal}(0, 2.5)$\
+$\beta_1 \sim \mbox{Normal}(0, 2.5)$\
+$\beta_2 \sim \mbox{Normal}(0, 2.5)$\
+$\beta_3 \sim \mbox{Normal}(0, 2.5)$\
+$\sigma \sim \mbox{Exponential}(1)$
 
-### Model justification
+### Model justification {#sec-justification}
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+A positive relationship is expected between batting average and number of wins as a higher batting average for a team signals a stronger offense which leads to runs.
+We also expect a positive relationship between defensive efficiency as a higher percentage indicates a higher prevention of runs and should lead to more wins.
+Lastly, a negative relationship between earned run average and wins is expected because a lower ERA means that the pitchers effectively prevented the opposite team from scoring runs.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+The model is ran using the *rstanarm* package (XX). The priors are assumed as rstanarm defaults. However, they are auto scaled based on the available 2000 - 2010 data.
 
@@
 
-Our results are summarized in .
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+mlb_model <- readRDS(here(file = "models/model.rds"))
+```
+
+The coefficients specified in the multiple regression model in @sec-setup are shown in @tbl-modelsum below. Our results are summarized in .
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-modelsum
+#| tbl-cap: Model of Team Wins per MLB Season based on Team Statistics
+
+modelsummary(list("auto-scaling priors" = mlb_model))
+
+```
 
@@
 Weaknesses and next steps should also be included.
+Weaknesses: irregular start for some teams, pitching clock
 

### 2024-04-21 scripts/01-download_data.R

@@ -11,4 +11,4 @@
 # Team data downloaded from https://www.baseball-reference.com/leagues/ and then 
 # converted from xls files to csv files to be saved in the raw data folder.
 # Hitting, pitching, and fielding data downloaded separately for each year to be combined later.
-# XX - XX data were downloaded, in total XX files.
+# 2000 - 2010 data were downloaded for model and 2024 data for prediction, in total 36 files.

### 2024-04-21 scripts/02-data_cleaning.R

@@ -71,10 +71,18 @@ b2010 <- read_csv("data/raw_data/2010B.csv")
 f2010 <- read_csv("data/raw_data/2010F.csv")
 p2010 <- read_csv("data/raw_data/2010P.csv")
 data2010 <- cbind(b2010[1:30, c("Tm", "BA")], f2010[1:30, c("DefEff")], p2010[1:30, c("ERA", "W")])
+#2024
+b2024 <- read_csv("data/raw_data/2024B.csv")
+f2024 <- read_csv("data/raw_data/2024F.csv")
+p2024 <- read_csv("data/raw_data/2024P.csv")
+data2024 <- cbind(b2024[1:30, c("Tm", "BA")], f2024[1:30, c("DefEff")], p2024[1:30, c("ERA", "W")])
 
-## Combining yearly data ##
+
+
+## Combining yearly data 2000 - 2010 ##
 mlb_10 <- rbind(data2000, data2001, data2002, data2003, data2004, data2005, 
                 data2006, data2007, data2008, data2009, data2010)
+
 
 ## Cleaning yearly data ##
 current <- c("Arizona Diamondbacks", "Atlanta Braves", "Baltimore Orioles", 
@@ -88,21 +96,29 @@ current <- c("Arizona Diamondbacks", "Atlanta Braves", "Baltimore Orioles",
              "Seattle Mariners", "St. Louis Cardinals", "Tampa Bay Rays", 
              "Texas Rangers", "Toronto Blue Jays", "Washington Nationals", 
              "Montreal Expos", "Tampa Bay Devil Rays",
-             "Los Angeles Angels of Anaheim", "Anaheim Angels")
+             "Los Angeles Angels of Anaheim", "Anaheim Angels", "Cleveland Guardians",
+             "Miami Marlins")
 change <- c("AZ", "ATL", "BAL", "BOS", "CHC", 
            "CWS", "CIN", "CLE", "COL", "DET", 
            "MIA", "HOU", "KC", "LAA", "LAD", 
            "MIL", "MIN", "NYM", "NYY", "OAK", 
            "PHI", "PIT", "SD", "SF", "SEA",
            "STL", "TB", "TEX", "TOR", "WSH", 
-           "WSH", "TB", "LAA", "LAA")
-mlb_10 <- mlb_10 |>
+           "WSH", "TB", "LAA", "LAA", "CLE",
+           "MIA")
+
[…]

### 2024-04-21 scripts/03-test_data.R

@@ -5,6 +5,8 @@
 # Contact: ycart.yang@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 02-data_cleaning.R
+# mlb_2024 data tests are less extensive as it is one year of the same format data 
+# cleaned the same way as mlb_data
 
 
 #### Workspace setup ####
@@ -14,28 +16,39 @@ library(arrow)
 
 #### Importing ####
 mlb_test <- read_parquet("data/analysis_data/mlb_data.parquet")
+mlb_2024_test <- read_parquet("data/analysis_data/mlb_2024.parquet")
 
 
 #### Test data ####
 
 ## Test Team ##
-mlb_test$Team |> unique() |> length() == 30
-mlb_test$Team |> is.character()
+mlb_test$Tm |> unique() |> length() == 30
+mlb_test$Tm |> is.character()
+mlb_2024_test$Tm |> unique() |> length() == 30
+mlb_2024_test$Tm |> is.character()
 
 ## Test AVG ##
-mlb_test$AVG |> is.numeric()
-mlb_test$AVG |> min() >= 0
-mlb_test$AVG |> max() <= 0.99
+mlb_test$BA |> is.numeric()
+mlb_test$BA |> min() >= 0
+mlb_test$BA |> max() <= 0.99
+mlb_2024_test$BA |> is.numeric()
+mlb_2024_test$BA |> min() >= 0
 
 ## Test ERA ##
 mlb_test$ERA |> is.numeric()
 mlb_test$ERA |> min() >= 0
+mlb_2024_test$ERA |> is.numeric()
+mlb_2024_test$ERA |> min() >= 0
 
 ## Test DefEff ##
 mlb_test$DefEff |> is.numeric()
 mlb_test$DefEff |> min >= 0
 mlb_test$DefEff |> max() <= 1
+mlb_2024_test$DefEff |> is.numeric()
+mlb_2024_test$DefEff |> min() >= 0
 
 ## Test W ##
 mlb_test$W |> is.numeric()
 mlb_test$W |> length() == 330
+mlb_2024_test$W |> is.numeric()
+mlb_2024_test$W |> length() == 30

### 2024-04-21 scripts/04-model.R

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models the relationships between three predictor variables on MLB team wins in a season.
+# Author: Tracy Yang
+# Date: 20 April, 2024
+# Contact: ycart.yang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 02-data_cleaning
 
 
 #### Workspace setup ####
@@ -13,25 +12,25 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+mlb_data <- read_parquet("data/analysis_data/mlb_data.parquet")
 
 ### Model data ####
-first_model <-
+model <-
   stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
+    formula = W ~ BA + DefEff + ERA,
+    data = mlb_data,
     family = gaussian(),
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+    seed = 11
   )
 
 
 #### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  model,
+  file = "models/model.rds"
 )
 
 

### 2024-04-21 paper/datasheet_template.qmd
Completed paper and other elements.

---
title: "Datasheet for 'mlb_data'"
author: 
  - Tracy Yang
thanks: "Code and data are available at: https://github.com/YcartXin/MLB_Team_Stats]."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "Using Major League Baseball key team statistics between 2000 and 2010, this paper investigates the outcome of the 2024 season using each team's current statistics on predicting their total wins out of 162 games. We investigate the common conception that baseball is a numbers game by evaluating the predictability of wins that a team would obtain during a seasion from looking at their batting average (BA), defensive efficiency ratio (DefEff), and earned run average (ERA). Despite rules changes in 2023 that may cause inaccuracies in the model, the prediction for current standings from a multiple linear regression model is quite accurate. The findings show that baseball indeed should consider statistics in assessing the competitiveness of a team."
format: typst
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to investigate MLB's team wins and whether it can be reasonably predicted by team statistics. We were unable to find a publicly available dataset in a structured format that had the variables for building the Baysian model to predict number of wins based on the chosen predictors.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Baseball Reference under Sports Reference LLD.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - TBD
4. *Any other comments?*
    - None.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	  - They represent the teams statistics of MLB teams throughout the years. 
2. *How many instances are there in total (of each type, if appropriate)?*
	- Three: Batting, Fielding, Pitching data.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- Instances from a larger set though it does not need to represent all MLB games, just reliably the season in 2024.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance consist of data related to batting, fielding, or pitching such as runs and total outs, etc.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- N/A
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- None from the years I have sampled.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- No.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- None.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- None.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset, once gathered, is self-contained.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- None.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- Depends on how much the viewer cares about baseball.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- None.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No.
16. *Any other comments?*
	- None.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data is directly observable although older data is taken from other sources.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The data collection process depends on both machines and humans such as the umpire calling outs and balls in baseball.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- Since the data is likely random, was sampling process is not needed.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Baseball enthusiasts and official MLB.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- Not sure the exact year and many past years are taken from other parties. However, data was definitely collected from 2000.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Not that I am aware of.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Third parties.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- No.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- I assume so as it is a public sport and players sign contracts.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- No.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No.
12. *Any other comments?*
	- None.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- No.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes, in the raw_data folder.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- Yes, MySQL.
4. *Any other comments?*
	- None.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- I am not sure.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No.
3. *What (other) tasks could the dataset be used for?*
	- Personal enjoyment.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- None.
6. *Any other comments?*
	- None.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- No.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- No.
3. *When will the dataset be distributed?*
	- N/A.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- N/A
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- None.
7. *Any other comments?*
	- None.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Baseball-Reference.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- Email address and phone.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- N/A.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Baseball-Reference provides a Question, Comment, Feadback, or Correction section on their website.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- No.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Yes. The data will exist on the website.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- They can contact Baseball-Reference through the feedback section.
8. *Any other comments?*
	- None.

\newpage

### 2024-04-21 paper/paper.qmd
Completed paper and other elements.

@@
 ---
-title: "MLB Team Statistics between 2000 - 2010 on Predicting Games Won per Season ^[Code and data are available at: https://github.com/YcartXin/MLB_Team_Stats]"
+title: "MLB Team Statistics Between 2000 to 2010 Build Baysian Model that Predicts Current Standing in the 2024 Season ^[Code and data are available at: https://github.com/YcartXin/MLB_Team_Stats]"
 author: 
@@
 abstract: "Using Major League Baseball key team statistics between 2000 and 2010, this paper investigates the outcome of the 2024 season using each team's current statistics on predicting their total wins out of 162 games.
-We also investigate the common conception that baseball is a numbers game by evaluating the predictability of wins that a team would obtain during a seasion from looking at their batting average (AVG), defensive efficiency ratio (DefEff), and earned run average (ERA)."
+We investigate the common conception that baseball is a numbers game by evaluating the predictability of wins that a team would obtain during a seasion from looking at their batting average (BA), defensive efficiency ratio (DefEff), and earned run average (ERA).
+Despite rules changes in 2023 that may cause inaccuracies in the model, the prediction for current standings from a multiple linear regression model is quite accurate.
+The findings show that baseball indeed should consider statistics in assessing the competitiveness of a team."
 format: 
@@
 mlb_data <- read_parquet(here("data/analysis_data/mlb_data.parquet"))
+mlb_2024 <- read_parquet(here("data/analysis_data/mlb_2024.parquet"))
 
@@
 Behind the entertaining personalities and eye-catching traditions of baseball is a game where teams and are meticulously crafted with calculation.
-Baseball's fascination with numbers persisted since its creation (Swhwarz, 2004).
-Runs scored have always been recorded to evaluate performance, while other statistics developed overtime into the complicated net of data that now reside upon the MLB official site (XX).
-The term "sabermetrics" even developed to describe in specific baseball statistics (XX) However,in Major League Baseball, a team's performance can differ greatly year to year.
+Baseball's fascination with numbers persisted since its creation @timeline.
+Runs scored have always been recorded to evaluate performance, while other statistics developed overtime into the complicated net of data that now reside upon the MLB official site @MLB The term "sabermetrics" even developed to describe in specific baseball statistics @pred However,in Major League Baseball, a team's performance can differ greatly year to year.
 A team's last year wins are not often indicative of the coming season.
@@
 
-Important statistics for a team in this paper is defined as the team's batting average (AVG), defensive efficiency ratio (DefEff), and earned run average (ERA).
+Important statistics for a team in this paper is identified as the team's batting average (BA), defensive efficiency ratio (DefEff), and earned run average (ERA).
 These three statistics each refer to an important aspect of a strong team: hitting, pitching, and fielding prospectively.
@@
 The specifics of these indicators will be elaborated on in @sec-data.
-Among present research, 2024 season predictions exist, however, they mainly focus on the win for a specific game or series for betting purposes, focusing on the player data.
+Among present research, 2024 season predictions exist, however, they mainly focus on the win for a specific game or series for betting purposes, focusing on the player data @pred.
 This paper is unique in its approach to predict and investigates the predictability of overall seasonal wins for baseball teams.
@@
 The rest of the paper is structured as follows.
-@sec-data (Data) will explain the source of the data used, its collection process, and the process of cleaning that led to the final data used.
-The data section will also explore the potential biases the collected and final data may possess.
-@sec-model...
+@sec-data (Data) explains the source of the data used, its collection process, and the process of cleaning that led to the final data sets used.
+The data section also explores the potential biases the collected and final data may possess.
+@sec-model (Model) explains and justifies the model used, then sets up the model using the 2000 - 2010 data.
+@sec-results (Results) will show the estimated values for the variables of interest and predict wins for the on-going season with the 2024 data.
+Lastly, @sec-discussion (Discussion) discusses the model coefficients, prediction results, weaknesses in the paper, as well as further possible exploration in the topic.
 
-The overall analysis shows...
+The overall analysis shows that team statistics in MLB is often indicative of the strength of the team and some top contenders in the league can be estimated by team statistics.
+Furthermore, BA seems to have the most effect on the number of wins a team should accumulate throughout a season among the three predictor variables.
+However, this investigation cannot validate the ability for current team statistics to predict future success in the same season.
 
@@
 
-The data used in this analysis of different statistics on MLB Teams are sourced from Sports Reference, more specifically, Baseball Reference ().
-The data sets used in this investigation focus on the team statistics of the 30 teams in MLB and span from 2000 to 2010.
+The data used in this analysis of different statistics on MLB Teams are sourced from Sports Reference, more specifically, Baseball Reference @BR.
+The data sets used for building the model in this investigation focus on the team statistics of the 30 teams in MLB and span from 2000 to 2010.
 The data sets were downloaded separately for each year among the batting, pitching, and fielding data.
 They were then combined into the final data set during the cleaning process.
-The analysis in this paper will be carried out in **R** @R using packages **tidyverse** @tidyverse, **dplyr** , **ggplot2** @ggplot2, **knitr** @knitr, **here** @here, **arrow** @arrow  XX.
+The prediction data of the 2024 season is downloaded in the same way and combined as a separate data set.
+The analysis in this paper will be carried out in **R** @R using packages **tidyverse** @tidyverse, **modelsummary** @modelsummary, **ggplot2** @ggplot2, **knitr** @knitr, **here** @here, **arrow** @arrow, **rstanarm** @rstanarm.
 
 ## Measurement and Collection
-The data available on Baseball Reference is gathered by many contributors through a variety of sources such as the Lahmen database, and much of older data is accessed from third-party websites such as Retrosheet ().
-Baseball Reference stores and manipulate the data gathered using MySQL ().
+The data available on Baseball Reference is provided by the official stats partner of the NBA, NHL, and MLB *Sportsradar* and gathered by many contributors through a variety of sources such as the Lahmen database @BR.
+Much of older data is accessed from third-party websites such as Retrosheet @BR.
+Baseball Reference stores and manipulate the data gathered using MySQL @BR.
 A potential weakness in the data collection includes missing data for certain teams, especially for early seasons in the 19th century when statistics were not kept in database and are easily lost.
@@
 
-There are in total 33 raw data sets of MLB team statistics in the raw data folder for constructing the model, these are split into 11 years and 3 categories (batting, pitching, fielding) for each team of every year from 2000 to 2010.
+There are in total 36 raw data sets of MLB team statistics in the raw data folder for constructing the model.
+The 33 for model building are split into 11 years and 3 categories (batting, pitching, fielding) for each team of every year from 2000 to 2010.
+The other three consist of the same categories but of 2024.
+Note that the statistics used for prediction in this paper is sourced on April 20th and is only representative of the statistics from games prior.
 Among each category, one representative indicator is chosen to demonstrate the overall strength of that category and be included as a variable for predicting wins over a season.
@@
 These variables include average runs per game or 2nd base reached over a season.
-Among these variables, we are interested in the batting average (**AVG**).
-AVG is calculated by total hits divided by total at-bats.
-Hits is defined by the number of times a batter strikes the ball into fair territory and reaches base, in the absence of an error or a fielder's choice ().
-AVG is a good indicator for offensive capabilities as it minimizes the impact of the pitcher by discounting walks and it is one of the universal tools to measure a hitter's success.
+Among these variables, we are interested in the batting average (**BA**).
+BA is calculated by total hits divided by total at-bats.
+Hits is defined by the number of times a batter strikes the ball into fair territory and reaches base, in the absence of an error or a fielder's choice @MLB.
+BA is a good indicator for offensive capabilities as it minimizes the impact of the pitcher by discounting walks and it is one of the universal tools to measure a hitter's success.
 
 The pitching data in a year contains 36 variables for each team detailing the bullpen's strength in initiating outs and preventing runs.
-Among variables in the data, we are interested in the Earned Run Average (**ERA**) calculated by 9 x earned runs/innings pitched ().
+Among variables in the data, we are interested in the Earned Run Average (**ERA**) calculated by 9 x earned runs/innings pitched @MLB.
 This number indicates the average earned runs allowed per 9 innings, which is better for a team if lower.
@@
 The fielding data for each year contains 16 variables that details the defensive abilities of a team to get the hitter out after a ball is in play.
-The variable chosen to represent this ability is Defensive Efficiency (**DefEff**), calculated by 1 - (H^[H = Hit] + ROE^[ROE = Reached on Error, when a batter reaches base due to a defensive error] - HR^[HR = Home Run]) / (PA^[PA = Plate Appearance] - BB^[BB = Base on Balls, better known as a walk which happens when the pitcher receives 4 balls from the umpire] - SO^[SO = Strike Out, when the batter is out of play from three strikes from the umpire] - HBP^[HBP = Hit By Pitch, when a batter is struck directly by a pitch without swinging and is awarded a walk] - HR), which is the percentage of balls in play converted into outs ().
+The variable chosen to represent this ability is Defensive Efficiency (**DefEff**), calculated by 1 - (H^[H = Hit] + ROE^[ROE = Reached on Error, when a batter reaches base due to a defensive error] - HR^[HR = Home Run]) / (PA^[PA = Plate Appearance] - BB^[BB = Base on Balls, better known as a walk which happens when the pitcher receives 4 balls from the umpire] - SO^[SO = Strike Out, when the batter is out of play from three strikes from the umpire] - HBP^[HBP = Hit By Pitch, when a batter is struck directly by a pitch without swinging and is awarded a walk] - HR), which is the percentage of balls in play converted into outs @MLB.
 This indicator estimates the ability of the catcher and fielders to efficiently prevent runs after a ball allows hitters to reach base and home.
@@
 
-The cleaned data now contains 330 observations of 5 variables:
+There are two cleaned data sets. The data set for model building contains 330 observations while the one for predicting contains 30. They both include 5 variables:
 
@@
 @tbl-sample shows a sample of the the cleaned data with 5 variables.
+
 ```{r}
@@
 #| tbl-cap: Sample of Cleaned MLB Prediction Data for Model Building
-mlb_data |> slice(1:8) |> kable()
+mlb_data |> slice(1:6) |> kable()
 
@@
 
-We can understand this data better by understanding the distribution of some of the variables.
+We can understand the model building data set better by understanding the distribution of some of the variables.
 For instance, we can understand the competitiveness of MLB by seeing the distribution of wins per season, whether the teams differ significantly by wins can help us analyze whether there is a significant disparity in the strengths of different teams.
@@
 
+Furthermore, we can see the relationship between the predictor variables by regressing the number of wins on each as shown in @fig-reg.
+
+```{r, fig.width = 2.8, fig.height = 2.3}
+#| echo: false
+#| fig-cap: Regressions of Number of Wins on each Predictor Variable
+#| label: fig-reg
+#| fig-subcap: ["Wins on BA", "Wins on DefEff", "Wins on ERA"]
+#| layout-ncol: 2
+
+mlb_data |>
+  ggplot(mapping = aes(x = BA, y = W)) +
+  geom_point(color = "steelblue", size = 1) +
+  theme_minimal() +
+  labs(x = "Batting Average", y = "Wins")
+
+mlb_data |>
+  ggplot(mapping = aes(x = DefEff, y = W)) +
+  geom_point(color = "steelblue", size = 1) +
+  theme_minimal() +
+  labs(x = "DefEff", y = "Wins")
+
+mlb_data |>
+  ggplot(mapping = aes(x = ERA, y = W)) +
+  geom_point(color = "steelblue", size = 1) +
+  theme_minimal() +
+  labs(x = "Earned Run Average", y = "Wins")
+
+```
+
+From @fig-reg, we can see that both batting average and defensive efficiency have positive relationships with the number of wins for a team.
+The directions for both graphs seem to be similar, even though DefEff has a more middle-concentrated data spread.
+This may suggest a stronger relationship for BA than DefEff.
+For earned run average, there is a clear negative trend which suggests that pitchers are of high importance in ensuring wins in baseball.
+However, this relationship is also expected as a high ERA for the team would mean that many points were scored against the team which is a more direct indication of losses than runs or good defensive plays.
+
 \newpage
@@
 
-A positive relationship is expected between batting average and number of wins as a higher batting average for a team signals a stronger offense which leads to runs.
+From @fig-reg, a positive linear relationship is expected between batting average and number of wins as a higher batting average for a team signals a stronger offense which leads to runs.
 We also expect a positive relationship between defensive efficiency as a higher percentage indicates a higher prevention of runs and should lead to more wins.
@@
 
-The model is ran using the *rstanarm* package (XX). The priors are assumed as rstanarm defaults. However, they are auto scaled based on the available 2000 - 2010 data.
+The model is ran using the **rstanarm** package. The priors are assumed as rstanarm defaults. However, they are auto scaled based on the available 2000 - 2010 data.
 
-# Results
+\newpage
+
+# Results {#sec-results}
 
@@
 
-The coefficients specified in the multiple regression model in @sec-setup are shown in @tbl-modelsum below. Our results are summarized in .
+The coefficient variables specified in the multiple regression model in @sec-setup are shown in @tbl-modelsum below.
 
@@
 
-modelsummary(list("auto-scaling priors" = mlb_model))
+modelsummary(list("posterior estimates" = mlb_model))
 
@@
 
-# Discussion
+The intercept is estimated in the model as -18.205. This does not make much sense as it indicates that a team with BA, DefEff, and ERA of 0 would have -18.205 wins which is not possible.
 
-## First discussion point {#sec-first-point}
+The 586.508 for $\beta_1$ is a much higher than the other beta variables. The stronger relationship is also suggested by @fig-reg where there exists a more consistent positive correlation.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The coefficient variable for DefEff is 24.2.
+Although the positive value is expected, the large difference between this coefficient and that of BA may suggest that offensive strength is more important than defensive strength for a MLB team.
 
-## Weaknesses and next steps
+$\beta_3$ is estimated to be around -16.557.
+The negative relationship is once again expected.
+However, since earned run average has a direct relationship with runs instead of hits from BA and DefEff, it would make sense to anticipate a lower value.
 
-Weaknesses and next steps should also be included.
-Weaknesses: irregular start for some teams, pitching clock
+```{r}
+#| echo: false
+#| label: tbl-pred
+#| tbl-cap: Top Five Teams for Predicted Wins of the 2024 Season
+
+pred <- c(mlb_model |> predict(newdata = mlb_2024) |> round())
+mlb_pred <- cbind(mlb_2024, pred) 
+mlb <- mlb_pred |> select(Tm, pred)
+rank <- mlb[order(-mlb$pred),] |> select(Tm, pred)
+rownames(rank) <- NULL
+rank |> slice(1:5) |> kable()
+```
+
+Using the model specified in @sec-setup, the current 2024 team statistics are used to predict the five teams with the highest number of wins in the league by the end of the season.
+@tbl-pred shows that the five teams are the Atlanta Braves, Milwaukee Brewers, Cleveland Guardians, New York Mets, and Kansas City Royals in this order.
+
+We can verify this prediction with the current (as of April 20th) MLB standings @BR.
+
+- The Atlanta Braves are currently tied with the New York Yankees at second with 14 wins
+- The Milwaukee Brewers is tied with Kansas City Royals and three other teams at third with 13 wins
+- The Cleveland Guardians leads the league at first with 15 wins
+- The New York Mets is tied with four other teams at fourth with 12 wins
+
+# Discussion {#sec-discussion}
+
+## Coefficient estimates
+
+From @tbl-modelsum, we see that batting average has a much higher effect on the number of wins than defensive efficiency or earned run average.
+This can be interpreted as the importance of offense showing through the statistics.
+For both DefEff and ERA, when successful, would prevent the opposing team form scoring runs.
+However, baseball is a game for which the team with the higher number of runs win.
+This means that strong defense is ineffectual without offense, and that a strong offensive team can at times mitigate lacking in fielding or pitching and result in wins.
+This effect is captured by the high coefficient estimate for BA in the model.
+Furthermore, frequent hits and runs puts mental pressure on the other team and effect their performance, a lead can often drive a game in a certain direction.
+
+On the other hand, as both DefEff and ERA captures the defensive abilities of a team, their coefficients would share the effects of keeping the other team at low or zero runs.
+These coefficients thus in comparison, more moderate than the batting average's.
+
+## Is baseball predictable by numbers {#sec-wins}
+
+The predictions made in @sec-results are not exactly the ranking of the MLB teams.
+However, the current league leader Cleveland Guardians is one of the predicted top fives total wins.
+The other teams predicted are also not far from the top.
+The least games won out of the five is New York Mets which has 12 wins and tied with four other teams at fourth.
+Considering that there are 30 teams and the rankings can drop quiet fast, the predicted teams clearly are comparatively strong teams in the league.
+
+We can look at the distribution of the predicted wins in @fig-wins2024 to further analyze the model.
+
+```{r}
+#| echo: false
+#| label: fig-wins2024
+#| fig-cap: Distribution of Wins of each Team in a MLB 2024 Season
+
+rank |>
+  ggplot(aes(x = pred)) +
+  geom_histogram(binwidth =  7, fill = "steelblue", alpha = 0.9, color = "white") +
+  theme_minimal() +
+  labs(
+    x = "Wins",
+    y = "Number of occurrences"
+  )
+
+```
+@fig-wins2024 differs from the collected distribution data from 2000 to 2010 in @fig-wins with a higher frequency distribution on both ends in comparison to the middle.
+@fig-wins resembles more like a Normal distribution although slightly skewed.
+@fig-wins2024 loses the resemblance to a Normal curve with a more even distribution.
+However, this is expected as there are only 30 data points in the predicted wins in comparison to the 330 that were collected.
+
+We can further check the reliability of the model by comparing the overall predicted number of wins in the league in comparison to the number of wins there should be in @tbl-total.
+
+```{r}
+#| echo: false
+#| label: tbl-total
+#| tbl-cap: Total Wins Predicted VS. Real Total Wins per Season
+total <- sum(rank$pred)
+real_total <- 81*30
+table <- data.frame(total, real_total)
+table |> kable()
+```
+
+Since whenever two teams play against each other there is always a winner, there should be a total number of wins a league should have by the end of the season.
+162/2*30 = 2430 The total wins predicted in the 2024 season is 2172.
+The model under-predicted the number of wins there should be in the league.
+However, looking at @fig-wins2024, we can see that the frequency of high wins is quite high.
+This means that the model severely under-estimated the wins of some teams that should be ranked more in the middle and slightly over-estimated some teams that should have less wins.
+Other the other hand, it may just be the fact that team statistics have not yet stabilized at a reasonable value with the limited number of games in the season.
+
+## Weaknesses
+
+There are a few weaknesses in the process of this investigation.
+Firstly, as we are using very early season data, there are many ties between teams in terms of number of wins and many teams have yet to display their true abilities.
+Each team has so far played around 20 games which is not enough to determine the direction for the rest of the season yet.
+Many players may have an irregular start in comparison to their last year statistics.
+The more data we have on the games each team plays, the team statistics would be updated and are more likely to be consistent with the accurate value.
+
+Furthermore, this analysis matched the currently standings to the predictions made by the current statistics in the league.
+Therefore, the teams with better statistics should naturally have more wins.
+The predictions made by the model indicates that team statistics in many ways is representative of the strengths of the team.
+However, it does not validate the ability that current baseball statistics can predict for wins for the whole season.
+
+Another weakness of this investigation is an important rule change implemented in MLB in 2023.
+The league added pitch timer that restricts the time that pitchers have in between pitches in order to increase the pace of play @MLB.
+This means that pitchers do not have the liberty to relax and prepare as much as they did before.
+This change may cause inaccuracies in the model by increasing/decreasing the estimate of ERA.
+However, there is not enough data since the rule change to build a model on so far.
+
+## Next Steps
+This analysis can be further developed by exploring the predictability of future wins based on a stable early season data.
+More specifically, analysis should be done by obtaining data after around 40 games for all teams and note down the predicted top winners in MLB at the end of the season to match the predictions made prior.
+By conducting such research over a few years, we would have a better idea of whether league winners show by statistics by a quarter of the season.
+
+Another way to explore MLB team predictions in depth is to consider the individual players statistics instead of team statistics.
+The team statistics likely overlooks the importance of a few individual key players and favour teams with an overall more even and higher batting, fielding, or pitching statistics.
+However, baseball games can be determined by one hit or defensive catch and individual statistics would better capture the true abilities of the team.
 
 \newpage
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In 1 we implement a posterior predictive check. This shows...
-
-In 2 we compare the posterior with the prior. This shows...
-
-## Diagnostics
-
-is a trace plot. It shows... This suggests...
-
- is a Rhat plot. It shows... This suggests...
-
-\newpage
-
-# References
